---
title: "Leaflet Practice - US Postsecondary Institutions"
author: "Armen Abnousi"
date: "July 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
setwd("~/coursera/coursera_developing_data_products/week2_pa/")
knitr::opts_chunk$set(echo = TRUE)
```
I have acquired college accreditation data from [here](https://ope.ed.gov/accreditation/dataFiles/Accreditation_2017_04.zip) and the US zipcode dataset from [here](https://boutell.com/zipcodes/zipcode.zip).
```{r, eval = TRUE, echo = FALSE, cache = TRUE}
download.file("https://ope.ed.gov/accreditation/dataFiles/Accreditation_2017_04.zip", destfile = "college_locations", method = "curl")
#download.file("https://ope.ed.gov/campussafety/api/dataFiles/file?fileName=Crime2016EXCEL.zip", destfile = "campus_safety", method = "curl")
download.file("https://boutell.com/zipcodes/zipcode.zip", destfile = "zip_to_latlang")
```
```{r, eval = TRUE, echo = FALSE}
cities <- read.csv(unzip(zipfile = "zip_to_latlang", files = "zipcode.csv"))
unis <- read.csv(unzip(zipfile = "college_locations", files = c("Accreditation_04_2017.csv")))
unis <- dplyr::select_(unis, "Institution_Name", "Institution_City", "Institution_State", "Institution_Zip")
unis$Institution_Zip <- sapply(as.character(unis$Institution_Zip), function(c){z <- unlist(strsplit(c, "-")); z[1]}, simplify = TRUE)
unis <- unique(unis)
unis$Institution_Zip <- as.integer(unis$Institution_Zip)
merged <- merge(unis, cities, by.x = c("Institution_Zip", "Institution_State", "Institution_City"), by.y = c("zip", "state", "city"))
colnames(merged)[colnames(merged) == "longitude"] <- "lng"
colnames(merged)[colnames(merged) == "latitude"] <- "lat"
```
There were initially `r dim(unis)[1]` accredited institutions in the dataset, by merging the data to find the institutions' latitude, longitude data, I could retrieve only `r dim(merged)[1]` of them. The map generated below shows these institutions and their locations. I do intend to add more options to this map later when I find the time! 

```{r, message = FALSE}
library(leaflet)
merged %>% leaflet() %>% addTiles() %>% addCircleMarkers(popup = merged$Institution_Name, clusterOptions = markerClusterOptions(), weight = 1) 
```